A Simple Type Theory for Category Theory

### 2022-02-08 paper/lipics-v2021-sample-article.tex
start working on kan extensions, fix errors

Next, let's prove synthetically that every functor $F : \varr{\cat
  C}{\cat D}$ is functorial in this notion of composition.
\begin{definition}
  For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
  : \homunary {\alpha_1} {\punittinXfromYtoZ {\cat C} {\alpha_1}
    {\alpha_2}} {\alpha_2} {\punittinXfromYtoZ {\cat D} {F \alpha_1}
    {F \alpha_2}}$ as
  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punittinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
  {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
  \]
  Note that
  \[ A[\id_\alpha/\alpha] = \id_{A} \]
  \[ A[f;g] = A[f];A[g] \]
\end{definition}
Note the similarity to the proof that all functions are congruences in
MLTT/HoTT.

@@

\subsection{Limits and Kan Extensions}

We can then use our notion of Hom of profunctors

The right Kan extension of $F$ along $J$ is the best (maximal) way to
fill in the following square:
\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
\begin{tikzcd}
C \arrow[d, "J"] \arrow[r, "F"] & D \\
C'                              &  
\end{tikzcd}\]
I.e. a functor $Ran_J F : C' \to D$ with a natural transformation
$Ran_J F \Rightarrow F$ that is universal something something.

@@

\begin{definition}
  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
  $J: \textrm{Functor}\,C'\,C$ be functors. Then the right Kan
  extension specification profunctor is a profunctor
  $\textrm{RanSpec}_J\,F : \textrm{Prof}\,D\,C$ defined as

>>>>

Next, let's prove synthetically that every functor $F : \varr{\cat
  C}{\cat D}$ is functorial in this notion of composition.
\begin{definition}
  For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
  : \homunary {\alpha_1} {\punitinXfromYtoZ {\cat C} {\alpha_1}
    {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat D} {F \alpha_1}
    {F \alpha_2}}$ as
  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punitinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
  {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
  \]
  Note that
  \[ A[\id_\alpha/\alpha] = \id_{A} \]
  \[ A[f;g] = A[f];A[g] \]
\end{definition}
Note the similarity to the proof that all functions are congruences in
MLTT/HoTT.

@@

\subsection{(co)-Limits and Kan Extensions}

As a further example of the synthetic approach to universal
properties, we consider \emph{Kan extensions}, which are a
generalization of limits, and their dual \emph{Kan lifts}, which
generalize co-limits.

The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
fill in the following square:
\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
\begin{tikzcd}
C \arrow[d, "J"] \arrow[r, "F"] & D \\
C'                              &  
\end{tikzcd}\]
I.e. a functor $Ran_J F : C' \to D$ with a natural transformation
$Ran_J F \Rightarrow F$ that is universal something something.

@@

\begin{definition}
  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
  $J: \textrm{Functor}\,C'\,C$ be functors. Then a right Kan extension
  is a functor $\textrm{Ran}_JF :\varr {\cat C'}{\cat D}$ that
  right-represents the following profunctor:
  \[ \]

  Then the right Kan
  extension specification profunctor is a profunctor
  $\textrm{RanSpec}_J\,F : \textrm{Prof}\,D\,C$ defined as
